---
title: "penguins-knn-mlr.Rmd"
author: "James Page"
date: "30/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook shows how to uild a kNN classifier that we can use to predict diabetes status from measurements of future patients. It is based on chapter 3 of [_Machine Learning with R, the tidyverse, and mlr_](https://www.manning.com/books/machine-learning-with-r-the-tidyverse-and-mlr) by Hefin Rhys (2020). 

This notebook shows how to use a kNN classifier to predict the sex of penguins using the ``palmerpenguins` data set - this is a set of observations of Antarctic penguins who live on the Palmer Archipelago. It is based on chapter 3 of [_Machine Learning with R, the tidyverse, and mlr_](https://www.manning.com/books/machine-learning-with-r-the-tidyverse-and-mlr) by Hefin Rhys (2020), and this article from Julia Silge on using the tidymodels: [Getting started with tidymodels and #TidyTuesday Palmer Penguins])(https://www.r-bloggers.com/getting-started-with-tidymodels-and-tidytuesday-palmer-penguins/)

First of all, lets have a look at the dataset:

```{r}
library(mlr)
library(tidyverse)
library(palmerpenguins)

penguinsTib <- as_tibble(penguins)
glimpse(penguinsTib)
summary(penguinsTib)
```

```{r}
ggplot(penguinsTib, 
       aes(bill_length_mm, flipper_length_mm, shape = species, color = species)) +
  geom_point() + 
  theme_bw()

ggplot(penguinsTib, 
       aes(bill_depth_mm, flipper_length_mm, shape = species, color = species)) +
  geom_point() + 
  theme_bw()

ggplot(penguinsTib, 
       aes(body_mass_g, flipper_length_mm, shape = species, color = species)) +
  geom_point() + 
  theme_bw()

ggplot(penguinsTib, 
       aes(body_mass_g, bill_length_mm, shape = species, color = species)) +
  geom_point() + 
  theme_bw()

ggplot(penguinsTib, 
       aes(body_mass_g, bill_depth_mm, shape = species, color = species)) +
  geom_point() + 
  theme_bw()
```

We'll create a new classifier task with the `makeClassifTask` function that we will use to predict the sex of a penguin based on bill_length_mm, flipper_length_mm and body_mass_g variables. In this example, sex is the target class, and bill_length_mm, flipper_length_mm and body_mass_g are the predictor classes.

```{r createTask}
# first we'll drop the colums we don't need - bill_depth_mm doesn't seem like it
# will be useful for us for instance

penguinsTib <- penguinsTib %>%
  filter(!is.na(bill_length_mm), !is.na(flipper_length_mm), !is.na(body_mass_g)) %>%
  select(species, bill_length_mm, flipper_length_mm, body_mass_g)

penguinsTask <- makeClassifTask(data = penguinsTib, target = "species")
```

Now we'll train a K-nearest neighbour classifier using the hyperparameter $k = 2$:

```{r createLearner}
knnLearner <- makeLearner("classif.knn", par.vals = list("k" = 2))
knnModel <- train(knnLearner, penguinsTask)
```

Let's use this model to make some predictions:

```{r}
knnPred <- predict(knnModel, newdata = penguinsTib)

performance(knnPred, measures = list(mmce, acc))
```

```{r}
penguinsTib
```


```{r}
newPenguins <- tibble(
  bill_length_mm = c(41.2, 44.9, 51.6, 50.2, 46.3),
  flipper_length_mm = c(192, 212, 203, 181, 200),
  body_mass_g = c(2990, 5310, 3475, 4470, 4560))

newPenguinsPred <- predict(knnModel, newdata = newPenguins)
newPenguins$species <- getPredictionResponse(newPenguinsPred)

newPenguins
```

```{r}
ggplot(newPenguins, 
       aes(bill_length_mm, flipper_length_mm, shape = species, color = species)) +
  geom_point() + 
  theme_bw()

ggplot(newPenguins, 
       aes(bill_length_mm, body_mass_g, shape = species, color = species)) +
  geom_point() + 
  theme_bw()

ggplot(newPenguins, 
       aes(body_mass_g, flipper_length_mm, shape = species, color = species)) +
  geom_point() + 
  theme_bw()
```

## Cross-Validation

Now lets run some cross-validation to test the value of our model. You'll recall that cross-validation is a way of evaluate the performance of your model on data it hasn’t seen yet by repeatedly sifting the data that is used for the training and test sets.

```{r include=FALSE}
kFold <- makeResampleDesc(method = "RepCV", 
                          folds = 10, reps = 50, 
                          stratify = TRUE)
kFoldCV <- resample(learner = knnLearner, task = penguinsTask, 
                    resampling = kFold, measures = list(mmce, acc))
```

```{r}
print(kFoldCV$aggr)
```

Lets look at the confusion matrix for one of the resamplings, and see how this compares with the one we generated previously:

```{r}
calculateConfusionMatrix(kFoldCV$pred, relative = TRUE)
```

## Tuning _k_

In this section we define a range of values for _k_ which can then be run as a parameter sweep. We'll use a grid search to try every single value in the parameter space.

```{r}
knnParamSpace <- makeParamSet(makeDiscreteParam("k", values = 1:10))
gridSearch <- makeTuneControlGrid()
```

We call the `tuneParams()` function to perform the tuning. We'll reuse the k-fold resampling we used previously.

```{r}
tunedK <- tuneParams("classif.knn", task = penguinsTask,
                     resampling = kFold,
                     par.set = knnParamSpace, control = gridSearch)
```

```{r}
knnTuningData <- generateHyperParsEffectData(tunedK)
plotHyperParsEffect(knnTuningData, 
                    x = "k", y = "mmce.test.mean",
                    plot.type = "line")
```
dy to train our final model, using a $k=`r tunedK$x`$ (rather than the $k=2$ we chose originally):

```{r}
tunedKnn <- setHyperPars(makeLearner("classif.knn"), par.vals = tunedK$x)
tunedKnnModel <- train(tunedKnn, penguinsTask)
```

Once we have our final model we can see how it will classify new patients. For example, say we have a set of new patients:

```{r}
newPenguins %>%
  mutate(species=NULL)

newPenguinsPred <- predict(tunedKnnModel, newdata = newPenguins)
newPenguins$species <- getPredictionResponse(newPenguinsPred)

newPenguins
```
